---
title: "R Notebook - Sorting"
output:
  html_notebook:
    theme: cosmo
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

# order
The function order() returns the index vector needed to sort the vector. This implies that sort(x) and x[order(x)] give the same result.

This can be useful for finding row numbers with certain properties such as "the row for the state with the smallest population". Remember that when we extract a variable from a data frame the order of the resulting vector is the same as the order of the rows of the data frame. So for example, the entries of the vector murders$state are ordered in the same way as the states if you go down the rows of murders.

**Instructions**

Now instead of the smallest population size, let's find out the row number, in the data frame murders, of the state with the smallest population size.

This time we need to use order() instead of sort().

Remember that the entries in the vector murders$population follow the order of the rows of murders.
```{r}
# Access population from the dataset and store it in pop
pop <- murders$population
# Use the command order to find the vector of indexes that order pop and store in object ord
ord <- order(pop)
# Find the index number of the entry with the smallest population size
ord[1]
```

```{r}
# Find the index of the smallest value for variable total 
which.min(murders$total)

# Find the index of the smallest value for population
which.min(murders$population)
```


# Using the output of order
Now we know how small the smallest state is and we know which row represents it. However, which state is it?

**Instructions**

Find the index of the smallest state using which.min(murders$population).

Define a variable states to hold the state names from the murders data frame.

Combine these to find the state name for the smallest state.
```{r}
# Define the variable i to be the index of the smallest state
i <- which.min(murders$population)

# Define variable states to hold the states
states <- murders$state
# Use the index you just defined to find the state with the smallest population
states[i]
```


# Ranks
You can create a data frame using the data.frame function. Here is a quick example:

temp <- c(35, 88, 42, 84, 81, 30)

city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

city_temps <- data.frame(name = city, temperature = temp)

**Instructions**

Define a variable states to be the state names from murders
Use rank(murders$population) to determine the population size rank (from smallest to biggest) of each state.
Save these ranks in an object called ranks.
Create a data frame with state names and their respective ranks. Call the data frame my_df.
```{r}
# Store temperatures in an object 
temp <- c(35, 88, 42, 84, 81, 30)

# Store city names in an object 
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

# Create data frame with city names and temperature 
city_temps <- data.frame(name = city, temperature = temp)

# Define a variable states to be the state names 
states <- murders$state

# Define a variable ranks to determine the population size ranks 
ranks <- rank(murders$population)
```


# Data Frames, Ranks and Orders
This exercise is somewhat more challenging. We are going to repeat the previous exercise but this time order my_df so that the states are ordered from least populous to most.

**Instructions**

Create variables states and ranks to store the state names and ranks by population size respectively.

Create an object ind that stores the indexes needed to order the population values, using the order command. For example we could define o <- order(murders$population)

Create a data frame with both variables following the correct order. Use the bracket operator [ to re-order each column in the data frame. For example, states[o] orders the abbreviations based by population size.

The columns of the data frame must be in the specific order and have the specific names: states, ranks.
```{r}
# Define a variable states to be the state names from the murders data frame
states <- murders$state

# Define a variable ranks to determine the population size ranks 
ranks <- rank(murders$population)

# Define a variable ind to store the indexes needed to order the population values
ind <- order(murders$population)

# Create a data frame my_df with the state name and its rank and ordered from least populous to most 
my_df <- data.frame(states=states[ind], ranks=ranks[ind])
my_df
```


# NA
The na_example dataset represents a series of counts. It is included in the dslabs package. You can quickly examine the object using

library(dslabs)

data

str(na_example)


However, when we compute the average we obtain an NA. You can see this by typing


mean(na_example)

**Instructions**

The is.na returns a logical vector that tells us which entries are NA. Assign the logical vector that is returned by is.na(na_example) to an object called ind.
Determine how many NAs na_example has, using the sum command.
```{r}
# Using new dataset 
library(dslabs)
data(na_example)

# Checking the structure 
str(na_example)

# Find out the mean of the entire dataset 
mean(na_example)

# Use is.na to create a logical index ind that tells which entries are NA
ind <- is.na(na_example)
# Determine how many NA ind has using the sum function
sum(ind)
```


# Removing NAs
We previously computed the average of na_example using mean(na_example) and obtain NA. This is because the function mean returns NA if it encounters at least one NA. A common operation is therefore removing the entries that are NA and after that perform operations on the rest.

**Instructions**

Write one line of code to compute the average, but only for the entries that are not NA making use of the ! operator before ind. 
```{r}
# Note what we can do with the ! operator
x <- c(1, 2, 3)
ind <- c(FALSE, TRUE, FALSE)
x[!ind]

# Create the ind vector
library(dslabs)
data(na_example)
ind <- is.na(na_example)
# We saw that this gives an NA
mean(na_example[!ind])
```

































